%---+----1----+----2----+----3----+----4----+----5----+----6----+----7----+----8----+----9----+----0----+----1


\documentclass[10pt, a4paper]{article}


\usepackage[spanish]{babel}
\hyphenation{
}

\usepackage{color}
\usepackage{graphicx}
\usepackage{epic}
\usepackage{curves}
\usepackage{multicol}
%\usepackage{array}
\usepackage{longtable}
\usepackage{amstext}
\usepackage{amsbsy}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{latexsym}  %\mathbb{R}
%\usepackage{wrapfig}
%\usepackage{multind}

\usepackage{graphics}
%\usepackage{psfig}
\usepackage{amsmath}
\usepackage[mathscr]{eucal}

\usepackage[latin1]{inputenc}

%\usepackage{array}
\usepackage{booktabs}



\usepackage{hyperref} 
\usepackage{xcolor}



%*************************************************************************************************************
%***  Formato del documento                                                                                ***
%*************************************************************************************************************

\textwidth       20 cm
\textheight      29 cm

\oddsidemargin   -2.3 cm
\evensidemargin   -2.5 cm 

\topmargin       -3.5 cm


\pagestyle{empty}

%\parskip = 1.5 \baselineskip


%*************************************************************************************************************
%***  Comandos para distintos conjuntos de numeros                                                         ***
%*************************************************************************************************************

\newcommand{\RR}{{\sf I}\hspace*{-1.5pt}{\sf R}}
\newcommand{\NN}{{\rm I}\hspace*{-1.5pt}{\rm N}}
\newcommand{\PP}{{\sf I}\hspace*{-1.5pt}{\sf P}}
\newcommand{\QQ}{\hspace*{2pt}{\sf I}\hspace*{-5pt}{\sf Q}}
\newcommand{\ZZ}{{\sf Z}\hspace*{-4.5pt}{\sf Z}}
\newcommand{\CC}{\hspace*{2.5pt}{\sf I}\hspace*{-5pt}{\rm C}}


%*************************************************************************************************************
%***  Comandos para subindices y superindices                                                              ***
%*************************************************************************************************************
 
\newcommand{\Ri}{_{_R}}
\newcommand{\Le}{_{_L}}



\newcommand{\repo}[2]
{  
\color{red}    
\href{#2}{#1}                     
\color{black}  
}




\newcommand{\week}[3]
{     
                    &    &         \\
      {\bf  #1 }  & #2 &#3    \\  \hline 
}


\newcommand{\online}[2]
{  
\color{red}    
\href{#2}{Online
 class #1}                     
\color{black}  
}


\newcommand{\onlinep}[2]
{  
\color{blue}    
\href{#2}{Online presentation #1}                     
\color{black}  
}

\newcommand{\onlinet}[2]
{  
\color{red}    
\href{#2}{Online class #1}                     
\color{black}  
}




\newcommand{\tabla}
{
\begin{center}
	
	\begin{tabular}{|p{0.5cm}|p{6cm}|p{12cm}|}
		\hline   
}


\newcommand{\etabla}
{

\end{tabular}
\end{center}


}



\begin{document}
%\setlength\extrarowheight{2cm}

%*************************************************************************************************************

\Large
\begin{center}                                          
	{\bf Space Systems. Master's degree. Advanced scientific programming techniques.  
	Numerical simulations of orbits. ETSIAE}                                                    
	{\bf Curso : 	2021-2022}   
                                                     
\end{center}

\normalsize
%\vspace{1.0cm}


\normalsize
\tabla 

\week{ \large{W} }
{ \Large{Numerical methods} }
{  \Large{Practice: NumericalHUB}
   \Large{Python \& Fortran}}


\week{1}
{
Presentation. Assignments: P1 and P2. 
Cauchy problem. Reduction to a first order system of equations. 

Kepler orbits. 
Euler and inverse Euler methods. Numerical example accomplished by a 
human computer. 
}
%----------------------------------------------
{ 
Visual Studio installation. 
Intel Fortran compiler installation. 
Python installation. 
NumericalHUB repository. Use the \texttt{MUSE\_orbits} module.
Numerical Kepler orbits with Euler and Inverse Euler methods. 
Implement milestone 2.
Vector languages. 
Use vector variables to write the Python code.
Check with different time steps. 

}



\week{2}
{
Global truncation error and local truncation error. 
Difference error equation. Upper bound for the global truncation error. 
Spectral radius.
 }
%----------------------------------------------
{
Abstractions grouped with modules. 
Explain the ingredients of the Cauchy problem. 
Arguments as functions. Describe the paradigm. 

Implement  milestone 2. 
%\online{3}{}

 Scope. Private, public, global and local variables and 
functions.
Paradigms: imperative versus declarative. 
Side effects. First class functions. Lexical closure.
Pure functions. 

}



\week{3}
{  
Error by means of Richardson extrapolation.  
Error convergence rate of a numerical method. 
 
}
%----------------------------------------------
{

Cauchy problem solver. Temporal scheme function.
Higher order functions. Abstraction hierarchical layers.   
Error functions based on Cauchy problem functions. 
Implement  milestone 3. 

Functional paradigm. 
Lambda functions. Lazy evaluation. 
Map, reduce and filter.
Layers of abstraction. 



}
 



\week{4}
{
Example:  $ \ddot  x + x =0$  and Euler 
 method.
 
%Kepler orbit in polar coordinates.  
%Phase space. 
%Poincaré maps.
}
%----------------------------------------------
{


\texttt{Cauchy\_ProblemS},    
 \texttt{Error\_Cauchy\_Problem} to estimate errors by Richardson extrapolation method. 
Implement  milestone 3.     

Top-down design  with functional composition. 
XTREME 
programming. TDD: Test Driven Development. 
}



\week{5}
{  
Region of absolute stability. Eigenvalues of the Jacobian.  
Characteristic polynomial of stability of a numerical scheme. 
Spectral mapping theorem.
}
%----------------------------------------------
{
  
\texttt{Temporal\_convergence\_rate} to evaluate 
the precision order of a temporal scheme. Implement  milestone 4. 


Continuation technique. 
Axiomatic design. ATP: Acceptance Test Protocols.
QuickCheck. Unit tests. 
Software design tools.

}


\week{6}
{
Fixed points or equilibrium points. Linearized system of equations. 
Jacobian at fixed points and eigenvalues.
Eigenvalues of the Jacobian. 
%Stability of linear ODEs with constant coefficients.
Stability of finite difference equations with constant coefficients. 
%Stability of periodic solutions. Monodromic matrix. Floquet multipliers.
Example:  $ \ddot  x + x =0$  and Crank Nicolson  method.
Region of absolute stability.
 

%\onlinet{5T}
}
%----------------------------------------------
{

 Implement  milestone 4. 
 
%Phase space of Kepler orbits.
%\online{4}{https://www.dropbox.com/s/u0ttfvsl0i3drpo/20201001.mp4?dl=0} 
%Tutorials of P1 assignment. 

Recursion in declarative languages versus iteration in imperative languages. 
Imperative example versus a declarative example: Newton method.  

Algorithm to draw regions of absolute stability. 

Numerical procedure to calculate the Jacobian. 
Stability regions.
Eigenvalues of the Jacobian at fixed points. Implement  milestone 5.  
}



\week{7}
{
Pointers and memory manage. N body problem. 
 }
%---------------------------------------------- 
{
%Tutorials of assignment P1. 

Examples of pointers: scalars, vectors and arrays of different rank. 
Simulate the N body problem with Python and Fortran.
}







\week{8}
{
Example Runge-Kutta of order 2 and geometrical interpretation. 
Explicit Runge--Kutta methods. Embedded Runge--Kutta methods. Computational effort. 
Algorithm to select a variable time step. 
}
%----------------------------------------------
{
Object example in Fortran and Python.
Circular restricted three body problem. 
Determination of the Lagrange points.  Stability of the Lagrange points.
Disturbances around the Lagrange points.
Implement  milestone 5. 
%\online{5P}




}



\week{9}
{

Coefficient determination of Runge--Kutta second order schemes. 
Explain polymorphism to integrate complex systems of differential equations. 
 %\onlinet{6T}{}
%\onlinet{10T}{https://www.dropbox.com/s/6dx9kwa54qp6k3o/20201112_T.mp4?dl=0}
}
%----------------------------------------------
{
Lyapunov orbits around $L_1, L_2, L_3 $. Implement  milestone 6.

Mapping in Fortran and Python.
Use \texttt{scipy} python library to 
 validate the results.  
  
}


\etabla



\tabla 

\week{\large{W}}
{ \Large{Numerical methods} }
{  \Large{Practice: NumericalHUB} 
   \Large{Python \& Fortran}}





\week{10}
{
Gragg-Burlisch-Stoer methods.
Algorithm to select a variable time step of GBS methods. 
 Computational effort. 
%\onlinet{7T}{}

 

%\onlinet{11T}
%{https://www.dropbox.com/s/n22jlnnpxm8zgzu/20201120_T.mp4?dl=0}
}
%----------------------------------------------
{

Arenstorf orbit. 
Shampine and Gordon orbits. Implement  milestone 6. 

%\online{6P}{} 
%\onlinet{11P}
%{https://www.dropbox.com/s/r4qwnlf6u3muabw/20201120_P.mp4?dl=0} 
Implementation of variable step algorithm in GBS methods. 

Comparison between GKS original implementation and modern Fortran implementation.
Implement  milestone 7.  
}


\week{11}
{
Explain why Leap-Frog is used in GBS. Explain filter to use Richardson extrapolation. 
Region of stability of Leap-Frog method. Example oscillator $ \dot z + i z = 0 
$ with $ z \in \CC $ simulated with Leap-frog scheme. Spurious solutions.

%\onlinet{12T}
%{}
 } %----------------------------------------------
{
 
%\onlinet{12P}
%{}




Object oriented programming. Classes and methods.
Encapsulation. Abstraction. Constructors.
Inheritance. Polymorphism.


}


\week{12}
{ 
Software design tool. Wrappers. Parallel programming. 
Coarrays in Fortran. Concurrent programming.
Existing software codes for the Cauchy problem.  
Python extensions and compilation. 
%\onlinet{8T}{}
%\onlinet{13T}
%{https://www.dropbox.com/s/3u9weiv00lll15s/20201203_T.mp4?dl=0}
}
%----------------------------------------------
{
Treatment of numerical exceptions in Fortran and Python. Try and except. Monads. 
Show examples of wrappers. Fortran wrappers in the NumericalHUB and scipy wrappers. 
Speed up. Speed up examples. Wrappers. Libraries in Fortran and Python. f2py. CPython.
Examples of parallel programming.

}



\week{13}
{ 
Create GitHub repository. Clone a remote repository. 
How to work with Git. How to collaborate. How to merge two branches. 
Merging from GitHub. Merging from Visual Studio. Revert specific commits. 
Create issues. GitHub classroom. 

 }
%----------------------------------------------  
{ 
Cases of functional programming success: Whatsapp (Erlang) and Facebook (Haskell). 



}





\week{14}
{Presentation of assignment P1 }
%---------------------------------------------- 
%\onlinep{P2}{https://www.dropbox.com/s/3zm2s60fzb9bxij/20201217_P2.mp4?dl=0} }
{Presentation of assignment P1 

}

\week{15}
{Presentation of assignment P1 }
%---------------------------------------------- 
%\onlinep{P2}{https://www.dropbox.com/s/3zm2s60fzb9bxij/20201217_P2.mp4?dl=0} }
{Presentation of assignment P1 
}

\etabla
\end{document}








\etabla



\vspace{1cm} 

\noindent A visual Srudio solution with a Python project and a Fortran project 
with a lot of different examples of advanced pprogramming techniques is hosted in: 

\begin{center} 
\repo{Software repository: Advanced-programming-Python-Fortran }
{https://github.com/jahrWork/Advanced-programming-Python-Fortran}
\end{center} 



\end{document}
































 Eigenvalues, Floquet 
multipliers, Lyapunov exponents.





\week{1}
{Presentation. Assignments: P1 (Lagrange points and stability) and P2 (Apolo 
13) }
{How to learn Applied Mathematics through modern FORTRAN. Use of the 
NumericalHUB}
{Python language. Code development.Software Development Life Cycle.}


\week{2}
{Cauchy problem. Euler method}
{First order ODE. Linear spring. API specification for the Cauchy problem. 
Kepler orbits with Euler method. }
{Python language. Functions, modules and classes. Vectors and matrices. NumPy 
and SciPy libraries.}



\week{3}
{Region of absolute stability. Richardson extrapolation. Convergence order of a 
numerical method. }
{Kepler orbits with Euler method. Different time steps. Relation with the 
region of absolute stability. Conservative versus dissipative methods. Error by 
means of the Richardson extrapolation. }
{Python language. Functions, modules and classes. Vectors and matrices. NumPy 
and SciPy libraries.}


\week{4}
{Embedded Runge--Kutta methods. Algorithm to select a variable time step. }
{Lagrange points of the three body problem. Jacobian matrix of the system. 
Eigenvalues of the Jacobian matrix. Stability of Lagrange points.   }
{Scope and pointers. API specification for the Cauchy problem. Implementation 
of the Euler method. Kepler orbit simulated with an Euler method and comparison 
with the NumericalHUB.  }


\week{5}
{Adams-Bashforth-Moulton methods. Milne method to select a variable time step.  
}
{Orbits with small disturbances around Lagrange points.}
{}






\week{2}{Vandermonde, Lagrange y Newton.}{Tabla de diferencias divididas. Error de truncamiento y error de redondeo.}
	
\week{3}{ Teorema $\pi (x)$. }
       {Acotación de $\pi (x)$ y de   $f^{(N + 1)} (\xi)$. \linebreak Acotación  de la función de Lebesgue.}
       
\week{4}{Acotación de $\pi (x)$ y de   $f^{(N + 1)} (\xi)$. \linebreak Acotación  de la función de Lebesgue.}
{Presentación hito 1.}        

\week{5}{Ceros y extremos de Chebyshev y su relación con puntos de la cuadratura de Gauss.}
{Presentación hito 2A. Corrección hito 1.}

\week{6}{Serie discreta y serie truncada. Plano físico y plano espectral y derivación en el plano espectral.}
{Presentación hito 2B. Corrección hito 2A.}

\week{7}{Interpolación continua a trozos. Error de truncamiento y redondeo de las fórmulas de diferencias finitas. \linebreak Fórmulas para las derivadas mediante diferencias finitas.}
          {Presentación hito 3. Corrección hito 2B.}

\week{8}{Fórmulas de diferencias finitas con tres puntos. $\Delta x $ óptimo.}
         {Presentación hito 4. Corrección hito 3.}

   
\week{9}{Problema de contorno 1D y 2D. }
{Presentación hito 5. Corrección hito 4.}
                                  

\week{10}{Problema de Cauchy en EDOS. Esquemas de discretización temporal. Métodos implícitos y explícitos. } 
          {Presentación hito 6. Corrección hito 5.}
   
\week{11}{Error local y error global de la discretización temporal. Ecuación del error y solución.}
{Presentación hito 7. Corrección hito 6.}   
                                 
\week{12}{Convergencia, consistencia y cero--estabilidad. Soluciones espúreas.}
{Presentación hito 8. Corrección hito 7.}

\week{13}{Acotación de la solución del error temporal. Radio espectral y relación con la región de estabilidad absoluta.
          Criterio de estabilidad.}
{Corrección hito 8.}


\week{14}{Problema de condiciones iniciales y de contorno. Método de las líneas. Ecuación del calor y ondas. }
         {Error espacial y error de truncamiento del operador espacial. Ecuaciones del error espacial y solución. }

\week{15}{Acotación de la solución del error espacial. Abscisa espectral y mal condicionamiento. 
    Región de estabilidad absoluta y criterio de estabilidad. }
{Ejemplos y discusión de estabilidad. Ensayo de examen tipo test. }

%\week{15}{}
 %                                   { }


\end{tabular}
\end{center}




\end{document}




